Dimensions of the impacts of land transitions in the Amazon and Cerrado

### 2023-02-09 manuscript/lct_se_ac.qmd
Add more content.

---
title: "Dimensions of the impacts of land transitions in the Amazon and Cerrado"
author: 
  - name: Hugo Tameirão Seixas
    email: seixas.hugo@protonmail.com
    url: hugotseixas.com
    affiliations:
      - name: State University of Campinas
format: 
  html:
    theme: lux
page-layout: article
bibliography: references.bib
link-citations: true
editor_options: 
  chunk_output_type: console
execute:
  freeze: auto
  cache: true
license: CC BY-NC-ND 4.0
---

```{r}
#| include: false

@@

Analysis of the drivers and impacts of native vegetation suppression in Brazil is not new. However, the approaches used to analyse this phenomena usually focus in few dimensions. I propose creating a modelling framework, using modern techniques, to predict and explain native vegetation suppression and its impacts on social and environmental variables. The study will also analyse the Amazon and Cerrado at the same time, and will explore whether predictions from one biome are related to predictions from the other biome.

@@

@@

To achieve the proposed goal, I will divide the work into three major parts:

@@

```{r}
#| echo: false

@@

The literature review will provide a theoretical and practical basis for the modelling section.

@@

```{r}
#| echo: false

@@

```{r}
#| echo: false
#| message: false

@@

@@

```{r}
#| echo: false

@@

@@

>>>>

---
title: "Dimensions of the impacts of land transitions in the Amazon and Cerrado"
author: 
  - name: Hugo Tameirão Seixas
    email: seixas.hugo@protonmail.com
    url: hugotseixas.com
    affiliations:
      - name: State University of Campinas
format: 
  html:
    theme: lux
page-layout: article
bibliography: references.bib
link-citations: true
editor_options: 
  chunk_output_type: console
execute:
  freeze: auto
  cache: true
license: CC BY-NC-ND 4.0
self-contained: true
---

```{r}
#| label: libraries
#| include: false

@@

Analyzing the drivers and impacts of native vegetation suppression is a complex task. The amount of factors influencing both phenomena can be numerous and heterogeneous. Existing literature suggests that native vegetation suppression can be caused by economic fluctuations [@hargrave-2012], conservation policies [@pfaff-2015], population distribution [@jusys-2016], land tenure [@cattaneo-2001], livestock and crop productions [@bowman-2016]. By the other side, vegetation removed is believed to impact human health [@santos-2018], climate regime [@silva-2023], regional hydrology [@cavalcante-2019], soil properties [@dias-2022], biodiversity [@gestich-2021].

Usually, the studies analyzing both phenomena focus in few variables, and normally struggle to relate their results to other areas of knowledge.

I propose creating a modelling framework, using modern techniques, to predict and explain native vegetation suppression and its impacts on social and environmental variables. The study will also analyse the Amazon and Cerrado at the same time, and will explore whether predictions from one biome are related to predictions from the other biome.

@@

An important aspect of the objectives, is that they are bound to the model behavior. This means that the results will not necessarily describe the phenomena in the real world, and this leap should be done with extreme care. This is true for any type of model developed to explain the reality.

@@

To achieve the proposed goal, I will divide the work into four major parts:

@@

```{r}
#| label: method-diagram
#| echo: false

@@

The literature review will provide a theoretical and practical basis for the modelling section. Information about what variables are being used, how are they being manipulated, what are the analysis techniques used to reach the results.

@@

```{r}
#| label: lit-review-diagram
#| echo: false

@@

```{r}
#| label: grid-map-figure
#| echo: false
#| message: false

@@

```{r}
#| label: database-diagram
#| echo: false
#| message: false

  graph [layout = dot, 
         overlap = false,
         rankdir = TB, 
         nodesep = 1.4, 
         ranksep = 0.8,
         tooltip = ' ']

  a [label = 'Data Source']
  b [label = 'Data Source']
  c [label = 'Data Source']
  d [label = 'Data Source']
  e [label = 'Pre Processing']
  f [label = 'Gridded Data']
  g [label = 'Database']

  a -> e b -> e
  c -> e d -> e
  e -> f f -> g

The data will be stored in a nested directories structured database, using give-style partitioning. The database will be composed by a collection of parquet files, a format with high compression, that facilitates processing large volumes of data.

@@

```{r}
#| label: model-table
#| echo: false

@@

Each model will follow the standard approach to develop supervised models.

```{r}
#| label: model-diagram
#| echo: false
#| message: false

grViz(
  height = 800,
  "digraph {

  graph [layout = dot, 
         overlap = false,
         rankdir = TD, 
         nodesep = 1.8, 
         ranksep = 0.5,
         tooltip = ' ']

  a [label = 'Database']
  b [label = 'Variables \n Selection']
  c [label = 'Data Split']
  d [label = 'Training Data']
  e [label = 'Validation \n Data']
  f [label = 'Test Data']
  g [label = 'Pre Processing']
  h [label = 'Initial \n Model Fit']
  i [label = 'Parameters \n Adjustments']
  j [label = 'Final \n Model Fit']
  k [label = 'Model \n Assessment']
  l [label = 'Model \n Interpretation']

  a -> b b -> c
  c -> d c -> e c -> f 
  d -> g g -> h 
  h -> i e -> i
  i -> j j -> k f -> k
  j -> l k -> l

In order to interpret the models, a series of methods will be used to explain the predictions, such as surrogate models, feature interactions, permutation feature importance, counterfactual explanations and Shapley values. These are all model agnostic methods, so they can be applied in any kind of regression or classification models.

@@

The idea of the project is to advance the understanding of why vegetation removals are happening, and than how are they impacting the environment and human well being. This have potential to shed light of what can be done to mitigate negative effects after the vegetation removal occurs.
